---
output: 
  word_document:
    reference_docx: memo_format.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(lubridate)
library(forecast)
library(RSocrata)
load('border_crossings.rda')
```

## FA49QRS

<br>

### MEMORANDUM FOR U.S. Army Analytic Agency
### ATTN: Ms. Janet Smith

<br>

### SUBJECT: United States Border Crossing Information Paper

<br>

<br>

1.Purpose: To provide insihts gathered from the border crossing/entry dataset.

<br>

2.Backgorund: The Border Crossing/Entry dataset contains `r nrow(data_pull)` records, pertaining to `r ncol(data_pull)` distinct measures (variables). The dataset specifically measures border crossing at the U.S.-Canada and the U.S.-Mexico border, at the port level.  Data area available for the following crossing methods: `r unique(data_pull$measure)`. Data is available from `r format(min(data_pull$date), format = '%Y')`, for `r data_pull %>% filter(., date <= '1997-01-01') %>% filter(!duplicated(port_name)) %>% select(port_name) %>% nrow()`  of `r data_pull %>%filter(!duplicated(port_name)) %>% select(port_name) %>% nrow()` port cities.

<br>

3.  Discussion: 

<br>

  a. First

<br>

  b. Second

<br>

  c. Third

<br>

4.  Findings:  This analysis does something…

<br>
